% !TEX root = ../main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

The disaggregated Keynesian cross of \citet{Teglio2025} shows how
heterogeneity in the marginal propensity to consume opens an output gap: when
income accrues to agents who do not spend it, demand leaks out of the
circular flow and the economy settles below capacity. The mechanism is a
demand mechanism, and the supply side plays no active role: capacity is given.

This paper asks what happens when the supply side is made endogenous in the
most direct way available --- by letting firms invest out of their own
profits and accumulate capital. The extension is not merely additive. Once
capital accumulates \emph{and} labour is hired endogenously, two channels
that point in opposite directions become active simultaneously:

\begin{itemize}
  \item \textbf{Supply.} Higher retention $\rho$ means more investment, more
  capital, and higher productive capacity. Taken alone, this raises output.
  \item \textbf{Demand.} In a demand-constrained regime, more capital means
  \emph{fewer workers are needed to serve the same demand}. Employment falls,
  the wage bill falls, and because workers have a higher propensity to consume
  than capitalists, demand falls with it.
\end{itemize}

Which channel dominates is the classical wage-led versus profit-led question
\citep{BhaduriMarglin1990,LavoieStockhammer2013}. Posed as a question about a
sign, it invites an answer in the form of a label. This paper's organising
finding is that the label is the wrong object: what the model determines is the
\emph{shape} of the response of output to retention, and the sign is a local
reading of that shape.

\paragraph{Three contributions.}

\textbf{First, $Y(\rho)$ is U-shaped, and the turn is where the economics is}
(\cref{sec:shape}). The response has a turning point $\rhostar$ that lies
inside the swept support and \emph{rises} with $\sigma$, from $0.42$ to $0.64$
across the range examined. The empirically anchored retention rate lies to the
\emph{left} of that turn for every $\sigma \geq 0.5$: the marginal effect of
retention at the point where the data put the economy is negative, while its
effect over a wide enough range is positive. Both are properties of one curve.
Summarised through a single slope, this structure presents as a sign frontier
$\sigstar \approx 0.65$ --- a reduction we retain, and declare as one.

\textbf{Second, the region in which retention depresses output is conditional
on fiscal institutions} (\cref{sec:government}). A balanced-budget
unemployment benefit, at a replacement rate inside the OECD band, eliminates
that region entirely, making retention expansionary at every elasticity
tested. The leakage of the unemployed from the circular flow is the engine of
the effect, and closing the leak removes it; a sign reported without its
institutional setting is therefore incomplete.

\textbf{Third, the wage curve is a source of fragility immune to demand
instruments} (\cref{sec:stress,sec:discussion}). Making wages flexible
destabilises the economy through a wage$\,\to\,U\,\to\,$capital-erosion
feedback, and three separate stabilisation hypotheses aimed at it --- slower
demand expectations, the unemployment benefit, and that benefit as a cushion
against firm heterogeneity --- are each falsified by the data designed to test
them, all three through that same channel.

A fourth contribution is methodological and kept deliberately off the main line
of argument: the sign depends on the estimator used to measure it, a two-point
chord and a whole-support slope disagreeing on $108$ identified design points
under a verdict rule fixed in advance (\cref{sec:estimator}). The same analysis
shows the conditional structure does not survive marginalisation --- viability
is governed almost entirely by the depreciation rate rather than by $\sigma$,
and at the rate the data imply the model does not exist at all
(\cref{sec:sa}).

\paragraph{A note on what this paper reports.}
An unusual proportion of what follows is negative. Three hypotheses are
falsified, one headline is shown not to generalise, and one measurement
instrument turns out to have been carrying part of its own answer. This is
deliberate. The project operates under a declared rule --- restated in
\cref{sec:method} --- that every number entering a table or a figure must
either be measured from a committed, reproducible driver or be labelled a
design target. The rule was adopted after an audit found that several
previously recorded ``results'' were in fact design targets that had never
been measured. Its consequence is that results contradicting the project's own
expectations are reported rather than recalibrated away; the negative findings
are the paper's substance, not its residue. Claims the project has withdrawn
under that rule are listed in \cref{app:retractions} rather than quietly
dropped.

The remainder is organised as follows. \Cref{sec:literature} places the model
in the literature and \cref{sec:model} presents it. \Cref{sec:calibration}
documents parameter anchoring and \cref{sec:method} the experimental protocol.
\Cref{sec:shape} to \cref{sec:sa} report the four blocks of results,
\cref{sec:discussion} draws the cross-cutting readings, \cref{sec:limits}
states the limitations, and \cref{sec:conclusion} concludes.
\Cref{app:validation} collects the validation record.
