% !TEX root = ../main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Relation to the literature
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Relation to the literature}
\label{sec:literature}

\paragraph{Agent-based and stock-flow-consistent macroeconomics.}
The model belongs to the tradition of macroeconomic ABMs that impose
stock-flow consistency on a disaggregated economy
\citep{GodleyLavoie2007,Caiani2016}, and shares with the Schumpeter-meeting-%
Keynes family \citep{Dosi2010} the ambition of letting aggregate regularities
emerge from interacting heterogeneous agents rather than from a
representative optimiser. It is deliberately smaller than those models: there
is one good, one price (the numéraire), no credit and no entry or exit. The
purpose is not to reproduce a full macroeconomy but to isolate a single
mechanism --- endogenous capital accumulation inside a demand-driven
economy --- with enough structure to make its sign an empirical question.

\paragraph{The elasticity of substitution.}
Comparing economies that differ in $\sigma$ is not straightforward: in the
textbook CES, varying $\sigma$ at fixed distribution and efficiency parameters
also moves the implied factor shares, so any behavioural difference is
unattributable. We therefore use the \emph{normalised} CES
\citep{KlumpDeLaGrandville2000,KlumpSaam2008,Klump2012}, which fixes a common
base point through which every $\sigma$-variant passes with the same factor
shares. We follow \citet{Temple2012} in reporting, rather than suppressing,
the caveat that normalisation is a comparison device and not an
identification strategy: it makes $\sigma$-variants commensurable, it does not
make $\sigma$ a deep parameter. \Cref{sec:frontier} accordingly reports the
sensitivity of our central estimate to the choice of anchor.

Empirically, the literature places $\sigma$ below unity: \citet{Chirinko2008}
weighs the evidence at $0.40$--$0.60$; \citet{ChirinkoMallick2017} estimate
$\approx 0.40$; the meta-regression of \citet{Knoblach2020}, covering 2{,}419
estimates from 77 studies, gives $0.45$--$0.87$ and rejects the
Cobb-Douglas hypothesis --- a rejection consistent with earlier direct tests
of the aggregate specification \citep{DuffyPapageorgiou2000,Antras2004}. A
dissenting macro literature requires $\sigma > 1$
to explain the global decline of the labour share
\citep{KarabarbounisNeiman2014}. We treat this disagreement as a reason to
\emph{sweep} $\sigma$ rather than choose it.

\paragraph{The wage curve.}
Wages are endogenised through the wage curve of
\citet{BlanchflowerOswald1994}, a \emph{level} relation between the local wage
and local unemployment with an elasticity near $-0.10$ ($\approx -0.07$ in the
meta-analysis of \citealp{NijkampPoot2005}). We use a level relation rather
than a Phillips-type change relation deliberately: without trend inflation, a
change relation produces perpetual wage drift at any $U \neq U^{\ast}$ and has
no steady state, which would make the comparative-statics design of this
paper impossible.

\paragraph{Distribution and effective demand.}
The question the model is built to answer is Kaleckian
\citep{Kalecki1971,BhaduriMarglin1990}: whether a redistribution toward
profits is expansionary or contractionary depends on whether the induced
investment exceeds the lost consumption. Our contribution is not a new theory
of that trade-off but a measurement of where its sign turns inside an
explicit, stock-flow-consistent, disaggregated economy --- and, more
importantly, a measurement of how little of that sign survives when the rest
of the parameter space is allowed to move.
