% !TEX root = ../main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The model
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The model}
\label{sec:model}

The economy is closed, single-good and fixed-price (the numéraire is~1). It
is populated by $n_f = 10$ firms and $N = 100$ households, of which a share
$\pi_{c} = 0.10$ are \emph{capitalists} who own firms and receive dividends;
the remainder supply labour. Time is discrete and one period is interpreted
as one year (\cref{sec:calibration}).

\subsection{Production: a normalised CES}

Each firm $i$ has a potential output given by a normalised CES in capital and
labour,
\begin{equation}
  \Ystar_i \;=\; Y_0 \Big[\, \pi_0 \big(K_i/K_0\big)^{r}
                    \;+\; (1-\pi_0)\big(L_i/L_0\big)^{r} \Big]^{1/r},
  \qquad r=\frac{\sigma-1}{\sigma},
  \label{eq:ces}
\end{equation}
with realised output rationed by demand,
$Y_i = \min\{\,D_i,\;\Ystar_i\,\}$. Here $(K_0,L_0)$ is the \emph{base point}
and $\pi_0$ the base-point capital share. Two properties matter.

\begin{enumerate}
  \item \textbf{Nesting.} \Cref{eq:ces} nests Cobb-Douglas at $\sigma = 1$
  and Leontief as $\sigma \to 0$. Because $Y_0$ is \emph{derived}
  ($Y_0 = A K_0^{\pi_0} L_0^{1-\pi_0}$) rather than measured, the $\sigma = 1$
  branch reproduces Cobb-Douglas exactly, for any base point.
  \item \textbf{Comparability.} $(K_0, L_0, \pi_0)$ are measured once, at
  $\sigma = 1$ and $\rho = 0.40$, and then \emph{frozen} across every
  experiment in the paper. Every $\sigma$-variant therefore passes through the
  same point with the same factor shares, so two economies differ \emph{only}
  by $\sigma$. This is a declared modelling choice, not an estimate.
\end{enumerate}

\subsection{The labour market: three limits}

Employment at firm $i$ is the minimum of three distinct constraints:
\begin{equation}
  L_i \;=\; \min\big\{\; L^{\mathrm{demand}}_i,\;\;
                        L^{\mathrm{profitmax}}_i,\;\;
                        N \;\big\},
  \label{eq:labour}
\end{equation}
where $L^{\mathrm{demand}}_i$ is the labour needed to serve expected demand,
$L^{\mathrm{profitmax}}_i$ is the point at which the marginal product of
labour equals the wage, and $N$ is the aggregate workforce. Which limit binds
identifies the regime: Keynesian (involuntary) unemployment, classical
unemployment, or full employment respectively.

\begin{declbox}[frametitle={Structural invariant: the workforce cap restores
diminishing returns}]
With a wage set outside the firm and an unbounded labour supply,
$L^{\mathrm{profitmax}}_i \propto K_i$, hence $\Ystar_i \propto K_i$: the
model becomes an $AK$ economy with constant returns to capital, unbounded
growth and no steady state. The cap $L \le N$ is therefore not a realism
detail but the structural feature that restores diminishing returns, and it
is asserted as such in the test suite.
\end{declbox}

Firms fire excess workers into an unemployed pool and fill vacancies by
random matching; the discreteness of hiring and the randomness of matching are
what give the model genuinely non-degenerate inter-seed variation.

\subsection{Wages: a Blanchflower--Oswald wage curve}

The wage is common to all firms and is set at the start of the period on the
\emph{previous} period's unemployment, which avoids a within-period
simultaneity between $w$ and $U$:
\begin{equation}
  w_t \;=\; \max\!\left\{\, w_{\min},\;
    \bar{w}\left(\frac{\max\{U_{t-1},\,U_{\min}\}}{U_{\mathrm{REF}}}
    \right)^{\!-\eta} \right\}.
  \label{eq:wagecurve}
\end{equation}
The elasticity $\eta$ is the parameter with empirical content;
$U_{\mathrm{REF}}$ is the unemployment level at which $w = \bar{w}$, measured
once at the anchor scenario and frozen; $U_{\min}$ and $w_{\min}$ are declared
conventions (guards against $w \to \infty$ at full employment and against a
deflationary spiral). At $\eta = 0$ the wage is constant at $\bar w$ and the
model reduces \emph{bit-for-bit} to its fixed-wage predecessor.

Profit is the residual, $\Pi_i = \text{sales}_i - w_t L_i$. There is no markup
parameter: with a fixed price and a parametric wage, distribution is
determined by $w_t$ alone, and the wage share is a \emph{measured outcome},
bounded above by the $\sigma$-dependent profit-max wage share.

\subsection{Investment and internal financing}

Firms finance investment entirely from retained earnings --- there is no
credit. Planned investment applies a retention ratio $\rho$ to last period's
profit, modulated by a capacity-utilisation accelerator:
\begin{align}
  \phi_t      &= \max\{0,\; 1 + \beta (u_{t-1} - \bar{u})\},
  \label{eq:accelerator}\\
  I^{p}_{t}   &= \clip\big(\rho\, \Pi_{t-1}\, \phi_t,\;
                  \underline{I},\; \Pi_{t}\big),
  \label{eq:investment}\\
  K_{t+1}     &= (1-\delta) K_t + I_t .
  \label{eq:capital}
\end{align}
The firm retains exactly what it invests and distributes the remainder as
dividends. Utilisation $u$ is measured against \emph{profit-max} capacity, not
against realised output: with $L$ chosen to serve expected demand, $\Ystar$
would otherwise track $Y$ by construction and the accelerator would receive a
dead signal.

\subsection{Households, government and stock-flow consistency}

Households consume out of income and wealth, subject to a liquidity
constraint,
\begin{equation}
  C_h = \min\big\{\, c_0 + c_1 y_h + \lambda a_h,\;\; a_h + y_h \,\big\},
  \label{eq:consumption}
\end{equation}
with workers' $c_1$ above capitalists'. An optional minimal government
(\cref{sec:government}) levies a flat tax on accrued income to fund an equal
transfer to the unemployed, indexed to the current wage, with the budget
balanced every period by construction; a replacement rate of zero reproduces
the no-government model bit-for-bit.

\begin{declbox}[frametitle={Stock-flow consistency, and where it is tested}]
The conserved quantity is
$\sum_h (\text{wealth}_h + \text{income}_h) + \sum_i \text{buffer}_i$, and it
is constant to within $10^{-9}$. The firm cash account is an
\emph{intra-period pass-through} that returns to zero every period: an earlier
design in which retained earnings accumulated as a stock between periods
sequestered money and produced a spurious demand spiral.

Crucially, the invariant is tested \emph{across the parameter range a sweep
will reach}, not only at the default configuration. \Cref{sec:ownership}
reports why: an ownership defect meant the invariant held at the default and
nowhere else, destroying up to 97\% of the money stock in 200 periods at
$\pi_c = 0.05$, while the tests --- run only at the default --- stayed green.
\end{declbox}

\subsection{Period sequence}

The order of events is load-bearing and is stated explicitly
(\cref{alg:sequence}). Employment is set \emph{before} households form demand,
because expected income depends on employment; investment settlement precedes
household settlement; and the government step sits between them so that the
tax falls on fully accrued income and the benefit arrives with the same
one-period lag as a wage.

\begin{algorithm}[H]
\caption{One period of the model}
\label{alg:sequence}
\begin{algorithmic}[1]
\State \textbf{wage determination:} set $w_t$ from \cref{eq:wagecurve} on
       $U_{t-1}$
\State \textbf{labour market:} firms plan employment via \cref{eq:labour};
       fire into pool; random matching fills vacancies
\State \textbf{demand formation:} households compute $C_h$
       (\cref{eq:consumption}); income is $w_t$ if employed, else $0$, plus
       dividends for capitalists
\State \textbf{investment plans:} firms apply \cref{eq:accelerator}--%
       \cref{eq:investment}
\State \textbf{demand registration:} consumption $+$ investment orders
\State \textbf{production:} $Y_i = \min\{D_i, \Ystar_i\}$; goods market
       rations; utilisation measured against profit-max capacity
\State \textbf{firm accounting:} wages paid, retention set aside, residual
       dividends declared
\State \textbf{investment settlement:} capital goods paid for,
       \cref{eq:capital} applied, residual returned as dividends
       (buffer $\to 0$)
\State \textbf{government:} balanced-budget tax and transfer (skipped at
       $rr = 0$)
\State \textbf{household settlement:} income credited, goods paid for
\end{algorithmic}
\end{algorithm}
