% !TEX root = ../main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Experimental protocol
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experimental protocol}
\label{sec:method}

\paragraph{Steady states, not histories.}
Unless stated otherwise, every reported figure is a mean over 20 random seeds
of the last 50 observations of a 2{,}000-step run. Runs are long because the
transient must be exhausted, not because 2{,}000 periods of history are
meaningful.

\paragraph{The sign frontier estimator.}
The quantity of interest is the sign of $\mathrm{d}Y/\mathrm{d}\rho$. We
estimate an OLS slope of $Y$ on $\rho$ over the \emph{common viable support}
(the set of $\rho$ values at which every configuration being compared
survives), and define $\sigstar$ as the value of $\sigma$ at which that slope
crosses zero. Confidence intervals come from a percentile bootstrap over
2{,}000 resamples \emph{of the seeds}. Restricting to a common support matters:
comparing slopes estimated on different supports would confound the effect
under study with differential survival.

\paragraph{Nesting checks as anti-drift sentinels.}
Every extension introduces a parameter whose null value must reproduce the
previous model exactly: $\sigma = 1$ (Cobb-Douglas), $\eta = 0$ (fixed wage),
$\lambda_e = 1$ (static expectations), $rr = 0$ (no government),
$\mathit{spread} = 0$ (homogeneous firms). Each is implemented as an explicit
branch and verified \emph{artifact-against-artifact} on committed panels. All
such checks passed with zero deviation at the time of their respective runs.

\begin{declbox}[frametitle={A methodological finding about the checks themselves}]
The criterion ``deviation $= 0.0$'' proved \emph{not reproducible over time}.
Code that had recorded 7/7 exact reproductions later deviated by 1 unit in the
last place (ULP) on the same cells. Eight hypotheses were excluded by
measurement (reporters, library versions, pandas reduction order, process
pool versus main process, performance versus efficiency cores, and others);
the cause remains unidentified. The magnitude across 160 cells and 24 metrics
is at most 2.1 ULP, does not amplify, and produces \emph{zero} regime flips,
so no economic conclusion moves --- and within a session, per-seed determinism
is intact. We report this rather than silently relaxing the criterion.

The replacement has since been implemented and the baseline re-fixed. It has
two limbs that cannot compensate for each other: a declared tolerance on
\emph{levels} (8 ULP with an absolute floor, roughly four times the measured
envelope), and a check on \emph{regime} indicators --- viability, which
constraint binds, the sign of every resolvable metric --- at tolerance exactly
zero, since that is where a drift would become scientific rather than
numerical. Two limits emerged while constructing it and are declared in the
source. A pure ULP criterion is unusable near zero: a metric legitimately
resting at zero registers thousands of ULP on a gap of $10^{-16}$, which is
why the absolute floor exists and why the regime limb declines to assert a
sign it cannot resolve. And the criterion must \emph{not} be applied to
differenced or fitted quantities --- a chord, a slope, a curvature --- because
catastrophic cancellation makes their relative error arbitrary even from
stable inputs; measured, the sensitivity analysis's slope deviates by 3410 ULP
while its own inputs deviate by 4. Those quantities are checked by sign
instead. Re-running the reference slice under the new criterion gives 7/7 with
a worst significant drift of $0.00$ ULP --- that is, it reproduces
\emph{exactly} today, where the earlier measurement found 2.1 ULP on the same
code. The drift is therefore intermittent, which is the strongest argument for
retiring exact equality: a criterion that passes or fails by the day carries
no information about what changed in the model.
\end{declbox}

\paragraph{Global sensitivity design.}
For \cref{sec:sa}, $\rho$ is a \emph{treatment} rather than a swept parameter:
each design point is evaluated at several values of $\rho$ under common random
numbers (the same seeds at each, differenced or regressed per seed, then
averaged). \Cref{sec:sa} uses two values and \cref{sec:sarepaired} four, the
four containing the two, so that the chord and the whole-support slope are
computed from the \emph{same runs} and their difference cannot be a different
sample. Without
common random numbers the quantity of interest is a difference of two noisy
numbers, and the sensitivity indices would decompose the variance of the seed
draw rather than of the model. Uniform ranges over 16 parameters are a
declared \emph{choice of ignorance}. Screening uses the elementary-effects
method of \citet{Morris1991} with a pruning rule frozen in the source before
any output existed; the survivors go to Sobol' estimation
\citep{Sobol2001,Saltelli2010} with bootstrap confidence intervals, computed
with SALib \citep{HermanUsher2017}. The number of seeds was fixed on evidence
by a pilot: the noise-to-signal ratio measured $0.207$, $0.170$ and $0.111$ at
3, 5 and 10 seeds, scaling as $1/\sqrt{n}$ as expected.

\paragraph{Implementation.}
The model is implemented in Python on the Mesa framework
\citep{Kazil2020}, with 512 automated tests covering stock-flow consistency
(parametrised over the ownership range), per-seed determinism, labour
accounting, every nesting limit, and directional regression pins. Each
committed data file has a committed generator script;
\cref{app:repro} gives the details, and \cref{app:notation} collects the
notation.
