% !TEX root = ../main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% An invariant that held only at the default
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{An invariant that held only at the default}
\label{sec:ownership}

This section reports a defect rather than a result, because the methodological
point outlasts the fix.

Firm ownership was assigned by cycling over \emph{households}: capitalist $i$
owned firm $i \bmod n_f$. That is a bijection \emph{only} at the default
configuration (100 households $\times$ 0.10 $=$ 10 capitalists, 10 firms). Off
the default it broke in both directions (\cref{tab:ownership}).

\begin{table}[htbp]
\centering
\small
\caption{The ownership defect, measured at seed 0 over 200 steps.}
\label{tab:ownership}
\begin{tabular}{lrrrr}
\toprule
$\pi_c$ & Capitalists & Firms with an owner & Money $t{=}0 \to t{=}200$
& $\sum$ net worth vs $K$\\
\midrule
0.05 & 5  & \textbf{5/10}  & $400.00 \to \mathbf{11.34}$ & $0.53\times$\\
0.08 & 8  & \textbf{8/10}  & $400.00 \to \mathbf{46.15}$ & $0.84\times$\\
0.10 & 10 & 10/10 & $400.00 \to 400.00$ & $1.05\times$\\
0.20 & 20 & 10/10 & $400.00 \to 400.00$ & $\mathbf{2.10\times}$\\
\bottomrule
\end{tabular}
\end{table}

Below the default, an ownerless firm's dividends and residual cash were paid
out inside a guard that tested for an owner, so they simply vanished --- a
direct violation of stock-flow consistency. Above it, the assignment
overwrote itself, leaving capitalists holding stale references whose capital
was still summed into net worth, inflating measured wealth inequality by a
factor of two.

\begin{declbox}[frametitle={The point worth keeping}]
The stock-flow invariant \emph{was} tested. It was tested at the default
configuration, it held there, and it held nowhere else --- and the test suite
was green throughout. This is exactly what a global sensitivity analysis
would have walked over in silence, dutifully reporting sensitivity indices for
a model that leaks money. An invariant tested at a point holds at that point.
Invariants must be parametrised over the range a sweep will reach
\emph{before} the sweep, not after.
\end{declbox}

The fix cycles over \emph{firms}, so every firm has exactly one owner for any
number of capitalists. At the default the assignment is unchanged, and a slice
of the committed panels spanning every dimension along which the code branches
reproduces exactly, so no previously reported result moves.
