% !TEX root = ../main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Discussion
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discussion}

\subsection{One channel, three falsified hypotheses}

The paper set out to measure a sign frontier and, in the process, found
something more robust than the frontier itself. Three separate stabilisation
hypotheses were posed and each was falsified:

\begin{enumerate}
  \item slower demand expectations would damp the collapse
  (\cref{sec:expectations}) --- they did not;
  \item a balanced-budget demand floor would shrink the collapse region
  (\cref{sec:government}) --- it \emph{enlarged} it, with the fiscal
  instrument saturated at its cap;
  \item that same benefit would cushion firm heterogeneity
  (\cref{sec:heterogeneity}) --- it \emph{lowered} the viability threshold.
\end{enumerate}

All three fail through the same channel and for the same reason. The
instability originates in the wage curve: unemployment moves the wage, the
wage moves profit-maximising employment (increasingly so as $\sigma$ rises),
employment moves unemployment, and each turn of the cycle leaves investment
short of depreciation, so capital erodes. Demand-side instruments act on the
wrong variable. Worse, in case (3) the benefit fails through a mechanism that
is fully traceable and slightly perverse: by \emph{succeeding} at lowering
unemployment, it causes the wage curve to raise the wage, which pushes the
weakest firm below its replacement investment.

The unifying boundary is worth stating plainly, because it is the most
transferable result in the paper: \textbf{demand instruments work where the
binding constraint is demand, and not against wage-driven capital erosion.}

\subsection{What the global analysis does and does not overturn}

The sensitivity analysis is destructive of the project's original headline and
should be read as such: over the defensible parameter space, wage-led is the
exception; $\sigma$ is nearly inert marginally; and viability is governed by a
parameter that the project had classified as a background convention. Any
version of this work that reported $\sigstar = 0.654$ as \emph{the} result
would have been reporting a cell as if it were a space.

What it does not overturn is the conditional finding itself. A conditional
frontier measured at defaults and a marginal effect averaged over fifteen
other parameters are different objects, and \cref{tab:sobol} gives positive
evidence that they should differ here. The honest summary is that the model
has a robust \emph{local} sign structure and a nearly inert \emph{global} one.

\subsection{The measurement instrument was part of the result}

\Cref{sec:estimator} reports the uncomfortable half of the reconciliation: a
material part of the gap between the conditional and marginal readings was our
own quantity of interest. A two-point chord and a whole-support slope are
different functionals of a curved response, and on the same committed data
they place $\sigstar$ at $0.447$ and $0.643$ with non-overlapping intervals.
Re-running the global analysis on the repaired quantity moves
$P(\text{wage-led} \mid \text{viable})$ from $0.095$ to $0.026$ --- and the
whole of that movement is traceable to $108$ individually identified design
points, those where the turning point of the U falls inside the chord's
window.

Two lessons generalise beyond this model. First, \textbf{a summary statistic
chosen before the shape of the response is known can encode a conclusion}: the
chord was fixed when $Y(\rho)$ was assumed monotone, and it kept returning
answers about a curve it could not represent. Second, the fix was not to pick
the better statistic but to \emph{stop compressing a curve into one number}.
That is why \cref{sec:shape} leads with the turning point and the side the
anchored range falls on, and treats the frontier as a declared reduction of
it: of the three summaries, only the position relative to the turn is
invariant to the analyst's choice of interval.

The episode has a plainer companion, recorded in \cref{app:retractions}: an
``inverted direction'' between the conditional and marginal analyses, which
this project reported as an open contradiction, did not exist in any
measurement --- it was an error in our own prose. The experiment designed to
explain it was still worth running, because it found a real defect on its other
axis. But the pairing is the lesson: a project's summary can contradict its own
tables, and nothing except re-deriving each claim from the measurements will
catch it.
