% !TEX root = ../main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Repairing the quantity of interest
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Repairing the quantity of interest}
\label{sec:bridge}

The tension reported in \cref{sec:inversion} is resolvable only by experiment,
and the experiment is designed as a $2\times2$: method (chord versus OLS)
crossed with conditioning (all other parameters fixed at their defaults versus
marginalised over the sensitivity ranges), on a shared grid of runs. The
verdict rule was fixed \emph{before} execution: a sign flip across the method
axis at fixed parameters implicates cause~(a); a flip across the conditioning
axis at constant method implicates cause~(b); both flipping means both
contribute and the weights must be quantified; neither flipping is a new
finding to be reported, not forced.

The experiment survived a discovery that might have dissolved it. Half of what
\cref{sec:inversion} set out to explain --- the apparent \emph{inversion} of
direction --- turned out to be a misreading of this project's own documents
rather than a property of the model, as the correction box there records. What
remained was a genuine disagreement between two estimators, which is what the
method axis was built to measure; and the conditioning axis, designed to test
a hypothesis about a phenomenon that did not exist, turned out to measure
something real in its own right. We report the design as it was fixed, not as
it would have been written knowing the answer.

The fixed-parameter arm costs no simulation: the committed panels already carry
the canonical $\sigma$ grid crossed with the canonical $\rho$ support at
twenty seeds, so the chord and the OLS slope are two linear functionals of the
same data and the same bootstrap delivers an interval for each. Only the
marginalised arm required new runs.

\subsection{\texorpdfstring{The 2$\times$2}{The 2x2}}

\Cref{tab:bridge} reports the completed design.

\begin{table}[htbp]
\centering
\small
\caption{Bridging experiment, complete. The fixed arm holds every parameter at
the configuration in which \cref{sec:frontier} identified $\sigstar$
($c_0 = 1.0$, $\eta = 0$); the marginalised arm draws the other fifteen
parameters from the \cref{sec:sa} ranges on the same $\sigma$ grid. All rows
use \emph{grid4}, the four-point $\rho$ support shared by both arms, so the
comparison is not confounded by the support.}
\label{tab:bridge}
\begin{tabular}{llrll}
\toprule
Cell & Parameters & Method & $\sigstar$ & 95\% CI\\
\midrule
1 & fixed        & chord & 0.447 & $[0.395,\,0.501]$\\
2 & fixed        & OLS   & 0.643 & $[0.596,\,0.701]$\\
3 & marginalised & chord & 0.938 & $[0.573,\,0.976]$\\
4 & marginalised & OLS   & \textbf{0.962} & $[0.723,\,0.985]$\\
\bottomrule
\end{tabular}
\end{table}

The mandatory anchoring control passes to fourteen significant figures: cell~2
evaluated on the full committed support returns
$\sigstar = 0.6540142777288407$ against the published $0.654$ with interval
$[0.616, 0.691]$. The bridge is therefore measuring the same object as
\cref{sec:frontier}, and any difference across the table is the treatment.

\paragraph{Verdict: cause (a).} At identical parameters, changing only the
estimator moves $\sigstar$ from $0.643$ to $0.447$ with \emph{non-overlapping}
intervals: for any $\sigma$ between those values the chord is negative while
the OLS slope is positive. The two statistics genuinely disagree in sign on the
same data, and they disagree about the statement the conditional result is
written in --- under OLS the empirical band $[0.40, 0.60]$ lies wholly on the
profit-led side, under the chord it straddles the frontier. By the rule fixed
in advance, that is cause~(a), and the quantity of interest is the defect.

\begin{declbox}[frametitle={What the rule does not capture, reported anyway}]
The rule asks whether the empirical band changes \emph{sides}. Marginalising
does not move it across --- cells 2 and 4 both place the band below $\sigstar$
--- so cause~(b) does not fire. But marginalising moves $\sigstar$ from $0.643$
to $0.962$, which is \emph{further} than the estimator moves it, and also with
non-overlapping intervals. It simply moves the frontier \emph{away} from the
band rather than across it.

Reporting only ``cause (a)'' would therefore let a threshold rule stand in for
the measurement. $\sigstar$ is strongly conditional on where the other fifteen
parameters are held; both displacements push the same way, and the empirically
supported band is profit-led more robustly than \cref{sec:frontier} alone
claimed. An earlier draft of this section predicted, from the fixed arm alone,
that conditioning ``must be doing the opposing work''. That inference was
correct: conditioning moves the chord frontier from $0.447$ to $0.938$.
\end{declbox}

\subsection{A quantity of interest suited to a U-shaped response}

The deeper repair is to stop summarising a U-shaped response with a single
slope. We therefore report the location of the turning point $\rhostar$ ---
the argument minimum of the quadratic fit --- alongside the slope, and, most
usefully, the \emph{side} on which the empirically anchored range of $\rho$
falls (\cref{tab:turning}).

\begin{table}[htbp]
\centering
\small
\caption{Turning point of $Y(\rho)$ and the position of the empirically
anchored range $\rho \in [0.35, 0.45]$ relative to it, at $\eta = 0$.}
\label{tab:turning}
\begin{tabular}{lrrrl}
\toprule
$\sigma$ & OLS slope & Chord & $\rhostar$ & Empirical $\rho$ lies\\
\midrule
0.30 & $+41.9$ & $+14.6$ & 0.420 & straddling\\
0.40 & $+26.8$ & $+5.2$  & 0.436 & straddling\\
0.50 & $+17.8$ & $-5.9$  & 0.462 & left of the turn\\
0.60 & $+5.6$  & $-20.7$ & 0.489 & left of the turn\\
0.70 & $-4.8$  & $-18.9$ & 0.513 & left of the turn\\
0.80 & $-11.7$ & $-33.3$ & 0.527 & left of the turn\\
1.00 & $-29.6$ & $-51.5$ & 0.569 & left of the turn\\
1.25 & $-45.3$ & $-68.8$ & 0.592 & left of the turn\\
1.50 & $-61.1$ & $-83.9$ & 0.640 & left of the turn\\
\bottomrule
\end{tabular}
\end{table}

\Cref{tab:turning} reconciles the two statistics mechanically. At
$\sigma = 0.5$ the turning point sits at $\rhostar = 0.462$, inside the
support: the OLS slope over $[0.35, 0.65]$ is positive because the rising
right branch is longer than the falling left branch, while the chord over
$[0.35, 0.55]$ is negative because it is weighted toward the left branch.
Neither is wrong; they are answers to different questions.

The economically interpretable statement is the third column. For every
$\sigma \geq 0.5$, the empirically anchored range of $\rho$ lies \emph{left of
the turning point} --- on the falling branch. The defensible formulation is
therefore not ``the economy is wage-led'' or ``profit-led'' but:

\begin{declbox}
\emph{Retention depresses output while $\rho$ remains below a
$\sigma$-dependent threshold $\rhostar$, and expands it beyond;
$\rhostar$ rises with $\sigma$, from $0.46$ at $\sigma = 0.5$ to $0.64$ at
$\sigma = 1.5$; and the empirically anchored retention rate lies below
$\rhostar$ for every $\sigma$ at or above $0.5$.}
\end{declbox}

This statement does not depend on which chord an analyst happens to pick, which
is exactly the property the original quantity of interest lacked.

\subsection{The sensitivity analysis, redone on the repaired quantity}
\label{sec:sarepaired}

\Cref{sec:sa} is then re-run end to end with the OLS slope over four $\rho$
values in place of the two-point chord. Parameters, ranges, frozen values and
the sampling seed are \emph{imported} from the original driver rather than
restated, so the quantity of interest is provably the only thing that differs;
and because the four-point support contains the two chord points, both
estimators are computed from the \emph{same runs}.

\paragraph{The screening changes.} Under the identical keep rule and seed,
$\mathtt{target\_utilization}$ enters the surviving set and
$\mathtt{benefit\_replacement\_rate}$ leaves it. A different quantity of
interest genuinely selects different parameters --- which is also why the
earlier runs could not have been recycled: a different survivor set is a
different design matrix.

\paragraph{The headline moves; the decomposition does not.}
\Cref{tab:repaired} puts the two estimators side by side.

\begin{table}[htbp]
\centering
\small
\caption{The repair, with both estimators computed on the same runs.
$P(\text{wl})$ abbreviates $P(\text{wage-led} \mid \text{viable})$.}
\label{tab:repaired}
\begin{tabular}{lrrr}
\toprule
Analysis & Fraction viable & $P(\text{wl})$ chord & $P(\text{wl})$ OLS\\
\midrule
\Cref{sec:sa}, chord only            & 0.483 & 0.095 & ---\\
Primary, $\sigma \in [0.40, 0.60]$   & 0.480 & 0.094 & \textbf{0.026}\\
Wide, $\sigma \in [0.30, 1.00]$      & 0.468 & 0.202 & \textbf{0.098}\\
\bottomrule
\end{tabular}
\end{table}

The chord column reproduces \cref{sec:sa} ($0.094$ against $0.095$; $0.202$
against $0.201$), the residual being the changed survivor set rather than the
estimator. The repaired quantity then cuts the wage-led fraction by a factor of
$3.6$ inside the empirical band, and by half in the wide check.

\paragraph{Why, to the point.} The difference is not diffuse. Chord and OLS
disagree in sign on $112$ of the $1596$ viable points; $110$ of those are
``chord negative, OLS positive'' and $2$ are the reverse, for a net of $108$.
The headline gap is $(0.0940 - 0.0263) \times 1596 = 108$ points. Every point
of the difference is individually identified, and what distinguishes them is
the mechanism predicted in \cref{tab:turning}: where the two estimators
disagree the mean turning point is $\rhostar = 0.473$, \emph{inside} the
chord's $[0.35, 0.55]$ window; where they agree it is $0.820$, outside the
support, where $Y(\rho)$ is monotone and any estimator sees the same thing.

\paragraph{The variance decomposition is robust to the repair.}
Total-order indices barely move: on the slope, $\delta$ goes from $0.966$ to
$0.900$, $\pi_0$ from $0.562$ to $0.561$, $\sigma$ from $0.024$ to $0.021$; on
viability, $\delta$ from $1.002$ to $0.916$. \emph{Which} parameters generate
the variance is robust to the estimator; the \emph{level} of the wage-led
probability is not. These are different questions, and the answer
\cref{sec:sa} gives to the first one stands.

\paragraph{The wide check, repeated.} \Cref{tab:widerepaired} is the direct
test of the reading error corrected in \cref{sec:inversion}.

\begin{table}[htbp]
\centering
\small
\caption{$P(\text{wage-led} \mid \text{viable})$ by $\sigma$ bin under both
estimators, on the same runs, $\sigma \in [0.30, 1.00]$.}
\label{tab:widerepaired}
\begin{tabular}{lrrr}
\toprule
$\sigma$ bin & $n$ viable & OLS & chord\\
\midrule
0.301--0.475 & 191 & 0.000 & 0.016\\
0.475--0.649 & 197 & 0.025 & 0.127\\
0.649--0.823 & 201 & 0.154 & 0.299\\
0.823--0.998 & 190 & 0.211 & 0.363\\
\bottomrule
\end{tabular}
\end{table}

Wage-led outcomes become \emph{more} common as $\sigma$ rises, under both
estimators --- the same direction as \cref{sec:frontier}, and never crossing
$0.5$ at any $\sigma$ tested. The tension reported as open in \cref{sec:sa} is
closed, and part of it was never there: only the estimator was genuinely at
issue, and its effect is to make the empirically supported band \emph{less}
wage-led, not more.

\paragraph{A limit on the turning point itself, declared.} In the marginalised
space $\rhostar$ falls inside the swept support in only $37.7\%$ of viable
points (median $0.323$; below the support in $58.7\%$ of cases), against 10 of
11 in the fixed-parameter cells of \cref{tab:turning}. Where the turn falls
outside, the U is \emph{not resolved there}, and it is reported as such rather
than extrapolated. The U-shaped statement above is a statement about the
conditional cells, and it is not silently promoted to the whole space.
