% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FRONT MATTER - title, abstract, contents
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\bfseries Endogenous Investment and Capital Accumulation\\[2pt]
       in a Disaggregated Keynesian Cross\\[6pt]
       \large A Normalised-CES Agent-Based Model with an Endogenous
       Labour Market}

\author{%
  Mattia \textsc{[Surname]}\thanks{%
  Correspondence: \texttt{[email]}. This paper documents an exploratory
  computational model. Every number reported here is either \emph{measured}
  from a committed, reproducible driver script, or explicitly declared as a
  design target or convention; the distinction is maintained throughout
  (\cref{sec:method}). Code, data and figures:
  \texttt{[repository URL]}.}\\[2pt]
  \normalsize [Institution]}

\date{\today}

\maketitle


\begin{abstract}
\noindent
We extend the disaggregated Keynesian cross of \citet{Teglio2025} with
endogenous investment and capital accumulation, so that both the demand and
the supply side of a heterogeneous-agent economy are endogenous. Production
is a \emph{normalised} CES \citep{KlumpDeLaGrandville2000} with elasticity of
substitution~$\sigma$; firms hire endogenously subject to three limits, the
wage follows a Blanchflower--Oswald wage curve, investment is financed
internally out of retained earnings, and the model is stock-flow consistent
across the whole parameter space. The central question is Kaleckian: does
higher retention raise output by building capacity (supply), or lower it by
displacing workers and shrinking the wage bill (demand)?

We find that the answer is governed by~$\sigma$, but that it is not well posed
as a single label. $Y(\rho)$ is \emph{U-shaped}, with a turning point
$\rhostar$ that rises with~$\sigma$ and lies inside the swept support in 19 of
22 conditional cells; the empirically anchored retention rate lies to the
\emph{left} of that turn for every $\sigma \geq 0.5$, so retention depresses
output there and expands it beyond. Read through a single slope, this
structure presents as a \emph{sign frontier} $\sigstar \approx 0.65$ --- a
description we retain, but as a local summary of a curve rather than as a
verdict on the economy. Three separate stabilisation hypotheses --- slower demand
expectations, a balanced-budget unemployment benefit, and that same benefit
as a cushion against firm heterogeneity --- are each \emph{falsified} by the
data they were designed to test, and all three fail through the same channel:
a wage$\,\to\,$unemployment$\,\to\,$capital-erosion feedback that originates
in the wage curve and is immune to demand-side instruments. This channel,
rather than the sign frontier, is the model's most robust finding.

A global sensitivity analysis over sixteen parameters then shows that the
headline does \emph{not} survive marginalisation: wage-led outcomes are the
exception, only $48\%$ of the empirically defensible space is viable at all,
and viability is governed almost entirely by the depreciation rate $\delta$
($S_T \approx 1.00$) rather than by~$\sigma$ ($S_T = 0.02$). At the
depreciation rate the data imply for this model's perimeter
($\delta \approx 0.09$), \emph{zero} of 843 sampled points survive --- a
structural limitation of a capital block without trend growth, not a parameter
to be tuned.

Finally we show that the quantity of interest was itself part of the result. A
$2\times2$ experiment, under a verdict rule fixed in advance, separates the two
candidate explanations of the gap between the conditional and marginal
readings, and returns the estimator: measuring a two-point chord across a
U-shaped response rather than a slope over its whole support moves $\sigstar$
from $0.643$ to $0.447$ on identical data, with non-overlapping intervals.
Re-running the entire sensitivity analysis on the repaired quantity cuts
$P(\text{wage-led} \mid \text{viable})$ from $0.095$ to $0.026$, while leaving
the variance decomposition almost unchanged --- \emph{which} parameters matter
is robust to the estimator, the \emph{level} of the headline is not. We also
document, and correct, a reading error of our own: an apparent inversion of
sign between the conditional and marginal analyses, which we had reported as
an open contradiction, was never present in the measurements.

\bigskip
\noindent\textbf{Keywords:} agent-based macroeconomics; stock-flow
consistency; elasticity of substitution; wage-led growth; global sensitivity
analysis; capital accumulation.

\smallskip
\noindent\textbf{JEL:} E11, E12, E22, C63, D31.
\end{abstract}

\clearpage
\tableofcontents
\clearpage
